Disparate views of disparity

### 2017-12-10 Analysis/disparity_methods.Rmd
Updated method proportions

```{r}
source("../Functions/fuzzy.sort.data.R")
source("../Functions/plot.data.R")
```

@@

```{r, fig.width = 18, fig.height = 6}
## Plotting the results
pdf("../Writing/Figures/MethodsProportions.pdf", width = 18, height = 6)
op <- par(mfrow = c(2, 3), bty = "n")
for(factor in 2:length(factors)) {
    plot.data(fuzzy.sort.data(data[,factor], factors[[factor]]), proportion = TRUE, ylim = c(0, 100), main = names(factors)[[factor]])
}
par(op)
dev.off()
```

@@

```{r}
## Removing the taxa column
multi_data <- apply(data[,-1], 2, as.character)

@@

## Transform the data as numeric
multi_data <- as.list(as.data.frame(multi_data))
multi_data <- lapply(multi_data, as.numeric)
multi_data <- matrix(unlist(multi_data), ncol = ncol(data[,-1]), byrow = FALSE)

@@

## Ordinating the distance matrix
ordinated_data <- stats::cmdscale(pairwise_data, k = sum(unlist(lapply(factors, length))[-1]) - 2)
#TG: Axis are selected as the number of factors - 2
```

@@

```{r, fig.width = 8, fig.height = 8}

@@

data <- cleaned_data$data

>>>>

```{r}
library(dispRity)
source("../Functions/fuzzy.sort.data.R")
source("../Functions/plot.data.R")
```

@@

```{r, fig.width = 12, fig.height = 12}
## Plotting the results
pdf("../Writing/Figures/MethodsProportions.pdf", width = 12, height = 12)
op <- par(mfrow = c(2, 2), bty = "n")
for(factor in c(2,4,5,6)) {
    plot.data(fuzzy.sort.data(data[,factor], factors[[factor]]), proportion = TRUE, ylim = c(0, 100), main = names(factors)[[factor]])
}
par(op)
dev.off()
```

@@

```{r}
## Removed columns
removed_factors <- c(1)

## Removing factors from the data
multi_data <- apply(data[,-removed_factors], 2, as.character)

@@

## Transform the data as numeric
multi_data <- as.list(as.data.frame(multi_data))
multi_data <- lapply(multi_data, as.numeric)
multi_data <- matrix(unlist(multi_data), ncol = ncol(data[,-removed_factors]), byrow = FALSE)

@@

## Ordinating the distance matrix
ordinated_data <- stats::cmdscale(pairwise_data, k = sum(unlist(lapply(factors[-removed_factors], length))) - 2)
rownames(ordinated_data) <- seq(1:nrow(data))
#TG: Axis are selected as the number of factors - 2
```

@@

```{r, fig.width = 8, fig.height = 8, eval = FALSE, echo = FALSE}

@@

## Get the group variables
Claddis_group_var <- list(factors$data_type[3], factors$morphospaces[2], factors$disparity[c(4,5)])
geomorph_group_var <- list(factors$data_type[c(1,2)], factors$morphospaces[1], factors$disparity[c(2,9)])
dtt_group_var <- list(factors$data_type[4], factors$morphospaces[4], factors$disparity[6])

select.group <- function(data, group) {
    ## Selecting the group variables
    multi.grep <- function(x, pattern, ignore.case = TRUE, ...) {
        ## Finding if the patten matches in x
        is_grep <- unlist(sapply(pattern, grep, x, ignore.case = ignore.case, simplify = FALSE, USE.NAMES = FALSE))
        ## Return true/false
        return(ifelse(length(is_grep) > 0, TRUE, FALSE))
    }

    data_type <- sapply(data[,2], multi.grep, pattern = group[[1]], USE.NAMES = FALSE)
    morphospace <- sapply(data[,4], multi.grep, pattern = group[[2]], USE.NAMES = FALSE)
    disparity <- sapply(data[,5], multi.grep, pattern = group[[3]], USE.NAMES = FALSE)

    all_match <- data_type & morphospace & disparity
    two_match_1 <-  morphospace & disparity
    two_match_2 <-  data_type & disparity
    two_match_3 <-  data_type & morphospace

    return(list("match3" = all_match, "match2" = list(two_match_1, two_match_2, two_match_3), "match1" = list(data_type, morphospace, disparity)))
}

## Select the groups
Claddis_groups <- select.group(data, Claddis_group_var)
geomorph_groups <- select.group(data, geomorph_group_var)
dtt_groups <- select.group(data, dtt_group_var)

## Select the full groups
Claddis_group <- Claddis_groups$match3
geomorph_group <- geomorph_groups$match3
dtt_group <- dtt_groups$match3
no_group <- !(Claddis_group | geomorph_group | dtt_group)

## Create the group variable
groups <- rep(NA, nrow(data))
groups <- ifelse(Claddis_group, "Claddis", groups)
groups <- ifelse(geomorph_group, "geomorph", groups)
groups <- ifelse(dtt_group, "dtt", groups)
groups <- ifelse(no_group, "none", groups)

## Get the groups IDs
Claddis_group_id <- which(Claddis_group)
geomorph_group_id <- which(geomorph_group)
dtt_group_id <- which(dtt_group)
no_group_id <- which(no_group)

## Plotting the group's

```{r, fig.width = 8, fig.height = 8}
col_palette <- c("red", "blue", "green3", "grey")

plot(ordinated_data[,1], ordinated_data[,2], xlab = paste0("PC1 (",percentages_scree[1], "%)"), ylab = paste0("PC2 (",percentages_scree[2], "%)"), pch = 19, col = col_palette[as.factor(groups)])

legend("bottomright", legend = levels(as.factor(groups)), col = col_palette, pch = 19, cex = 0.5)
```

## Testing the effect of groups

```{r}
## Creating the custom groups list
groups_list <- list("Claddis" = Claddis_group_id, "geomorph" = geomorph_group_id, "dtt" = dtt_group_id)

## Creating the groups for the pairwise matrix and the ordinated one
pair_matrix <- as.matrix(pairwise_data)
rownames(pair_matrix) <- seq(1:nrow(data))
groups_pair <- custom.subsets(pair_matrix, group = groups_list)
groups_ordi <- custom.subsets(ordinated_data, group = groups_list)

## Effect of groups
vegan::adonis2(pairwise_data ~ group, data = data.frame(matrix(groups, ncol = 1, dimnames = list(seq(1:nrow(data)), c("group")))))

## Effect of groups (individual)
groups_data <- data.frame(cbind(as.factor(Claddis_group), as.factor(geomorph_group), as.factor(dtt_group), as.factor(no_group)))
rownames(groups_data) <- seq(1:nrow(data))
colnames(groups_data) <- c("Claddis", "geomorph", "dtt", "nope")
vegan::adonis2(pairwise_data ~ Claddis + geomorph + dtt + nope, data = groups_data, permutations = 100)
